The effect of climate forcing on numerical simulations of the Cordilleran Ice Sheet at the Last Glacial Maximum

### 2013-12-09 reviews/initial-response.tex
Small corrections (Irina)

\opening{Dear Editor,}

@@

\begin{itemize}

@@

Thank you for raising this point. A justification was added in the manuscript. In this study, we focus mainly on glacier extent, on which geothermal heat flux is not the major control. However we will be happy to add a sensitivity experiment to the paper if this point is risen again at the discussion stage.

@@

\todo{Develop this point.}

@@

The LGM ice margin reconstruction visible on Figs.~1 and~15 was added to Fig.~6, which shows the extent of modelled ice cover for each climate forcing used. However we did not add it to Figs.~7 and 13 as we felt the resulting figures contained too much information (Fig.~a).

>>>>

\opening{Dear Frank Pattyn,}

@@

% ----------------------------------------------------------------------
\begin{itemize}

@@

Thank you for raising this point. A justification was added in the manuscript. In this study, we focus mainly on glacier extent, on which geothermal heat flux is not the major control. However we are currently working on an improved geothermal forcing which will be included in transient simulations of the Cordilleran Ice Sheet in the future.

@@

There exist different refreezing parametrizations in the literature. We developed this point in the manuscript. As we have not assessed which parametrization scheme, if any, would be appropriate for our set-up, we preferred to disregard refreezing rather than adding a new source of uncertainty in our model.

@@

The LGM ice margin reconstruction visible on Figs.~1 and~15 was added to Fig.~6, which shows the extent of modelled ice cover for each climate forcing used. However we did not add it to Figs.~7 and 13 as we felt that the resulting figures contained too much information (Fig.~a).

### 2013-12-09 sections/discuss.tex
Small corrections (Irina)

Furthermore, the surface mass balance model crudely simplifies of processes taking place at the glacier surface. Refreezing of melted snow and ice and retention in the snow pack was shown to alter surface mass balance of glaciers \citep{trabant-mayo-1985}, yet it is not accounted for in our PDD model. Based on scarce observations over the Greenland Ice Sheet, different parametrizations of this effect have been developped \citep{janssens-huybrechts-2000,fausto-etal-2009b}. However, these parametrizations were shown to disagree with more complex models \citep{reijmer-etal-2012}, and their portability to a different region and time period has not been proven. In addition, we simulate temperature variability by using a constant, uniform value of daily temperature standard deviation. This approach implies large biases of surface mass-balance \citep{charbit-etal-2013,rau-rogozhina-2013,seguinot-2013}, particularly over regions with such various climates as experienced within our model domain.

>>>>

Furthermore, the surface mass balance model crudely simplifies the processes taking place at the glacier surface. Refreezing of melted snow and ice and retention in the snow pack was shown to alter surface mass balance of glaciers \citep{trabant-mayo-1985}, yet it is not accounted for in our PDD model. Based on scarce observations over the Greenland Ice Sheet, different parametrizations of this effect have been developed \citep{janssens-huybrechts-2000,fausto-etal-2009b}. However, these parametrizations were shown to disagree with more complex models \citep{reijmer-etal-2012}, and their portability to a different region and time period has not been proven. In addition, we simulate temperature variability by using a constant, uniform value of daily temperature standard deviation. This approach implies large biases of surface mass-balance \citep{charbit-etal-2013,rau-rogozhina-2013,seguinot-2013}, particularly over regions with such various climates as experienced within our model domain.

### 2013-12-09 sections/model.tex
Small corrections (Irina)

The Shallow Shelf Approximation (SSA) is used as a ``sliding law'' for the Shallow Ice Approximation (SIA) \citep{bueler-brown-2009,winkelmann-etal-2011}. SIA and SSA velocities are computed by finite difference methods on a 10\,km-resolution horizontal grid of 300 by 150 points (the modelling domain). Ice softness depends on temperature and water content through an enthalpy formulation \citep{aschwanden-blatter-2009,aschwanden-etal-2012}. Enthalpy is computed in three dimensions in up to 51 irregularly spaced layers within the ice, and temperature is additionally computed in 11 regularly spaced layers in bedrock to a depth of 1\,km. Surface air temperature from the climate forcing provides the upper boundary condition to the ice enthalpy model, and a geothermal heat flux of 70\,\unit{mW\,m^{-2}} provides the lower boundary condition to the bedrock thermal model. Although this uniform value does not account for the high spatial geothermal variability of the region \citep{blackwell-richards-2004}, it is on average representative of available heat flow measurements \citep{artemieva-mooney-2001,blackwell-richards-2004}. In addition, whereas spatial variations of geothermal heat flux may significantly affect basal conditions of ice sheets \citep{pattyn-2010}, this study focuses on reconstructing the former extent of ice cover, on which it has relatively little effect \citep{rogozhina-etal-2012}.

>>>>

The Shallow Shelf Approximation (SSA) is used as a ``sliding law'' for the Shallow Ice Approximation (SIA) \citep{bueler-brown-2009,winkelmann-etal-2011}. SIA and SSA velocities are computed by finite difference methods on a 10\,km-resolution horizontal grid of 300 by 150 points (the modelling domain). Ice softness depends on temperature and water content through an enthalpy formulation \citep{aschwanden-blatter-2009,aschwanden-etal-2012}. Enthalpy is computed in three dimensions in up to 51 irregularly spaced layers within the ice, and temperature is additionally computed in 11 regularly spaced layers in bedrock to a depth of 1\,km. Surface air temperature from the climate forcing provides the upper boundary condition to the ice enthalpy model, and a geothermal heat flux of 70\,\unit{mW\,m^{-2}} provides the lower boundary condition to the bedrock thermal model. Although this uniform value does not account for the high spatial geothermal variability in the region, it is on average representative of available heat flow measurements \citep{artemieva-mooney-2001,blackwell-richards-2004}. Whereas spatial variations in geothermal heat flux may significantly affect basal conditions of ice sheets \citep{pattyn-2010}, this study focuses on reconstructing the former extent of ice cover, on which it has relatively little effect \citep{rogozhina-etal-2012}.
